% Workshop draft. Compile-once skeleton; refine as experiments stabilize.
\documentclass{article}
\usepackage[a4paper, margin=1in]{geometry}
\usepackage{amsmath, amssymb}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{cite}
\usepackage{algorithm}
\usepackage{algorithmic}

\title{Calibration-Aware Adaptive Aggregation\\
       for Foundation Time-Series Forecasters}
\author{Anonymous}
\date{}

\begin{document}
\maketitle

\begin{abstract}
Foundation time-series forecasters (Chronos, Moirai, TimesFM) achieve strong
point accuracy on standard benchmarks but exhibit systematic miscalibration
of their predictive intervals---e.g.\ Chronos-Bolt-base reports an 80\%
interval empirically covering $0.69$--$0.88$ of test targets across the
ETT-small suite. We study whether a learned per-window orchestrator over a
heterogeneous panel of forecasters can recover near-nominal coverage at
near-best-single CRPS, comparing directly to split-conformal prediction
applied to the best single agent. We propose \emph{Sequential Refinement}:
an RL-trained policy that, conditioned on input features and the current
ensemble's running quantile state, decides at each step whether to halt
or invoke an additional forecaster. Empirically, split-conformal
\emph{overcorrects} on $3/4$ ETT-small datasets, inflating intervals to
$0.95$--$0.97$ coverage because val and test exhibit a non-trivial
distribution shift that violates conformal's exchangeability assumption.
The Sequential Refinement orchestrator, in contrast, attains mean coverage
deviation $|cov_{80} - 0.80| = 0.022$ (vs.\ $0.093$ for the best-conformal
agent) and beats conformal-calibrated best-single CRPS on ETTh1 while
remaining competitive on the others. We also report a same-distribution
diagnostic that bounds the method's headroom at $\approx 6\%$ over
best-single CRPS, characterizing the regime where multi-agent routing
can and cannot improve over post-hoc per-agent calibration.
\end{abstract}

\section{Introduction}
\label{sec:intro}

% [Motivation]
Recent foundation models for time-series forecasting---Chronos~\cite{ansari2024chronos},
Moirai~\cite{liu2024moirai}, TimesFM~\cite{das2024timesfm}---deliver
strong zero-shot point accuracy across standard benchmarks. However,
their predictive intervals are often miscalibrated: across the ETT-small
suite, the empirical coverage of the model's 80\% interval ranges from
$0.65$ (Chronos-base on ETTm1) to $0.88$ (Chronos-base on ETTm2),
deviating from the nominal $0.80$ by $0.05$--$0.15$ in both directions.

% [Gap]
Calibration matters whenever downstream decisions consume the predictive
distribution (risk pricing, anomaly detection, capacity planning). Two
families of approaches address this: \emph{post-hoc} methods such as
conformal prediction~\cite{vovk2005algorithmic,romano2019cqr}, which
adjust a single model's intervals using held-out conformity scores; and
\emph{multi-model} aggregation, which combines distributions from a
panel of forecasters. We argue that the multi-model approach has a
specific structural advantage---decorrelation of miscalibration
directions---that post-hoc per-model calibration cannot exploit.

% [Contribution]
We propose Sequential Refinement: an RL-orchestrated protocol that
adaptively invokes forecasters from a heterogeneous panel
\{ARIMA, Chronos-Bolt-base, Chronos-Bolt-tiny, Moirai\} per input window.
The policy observes a window-features vector, the current ensemble's
quantile state, and a cost budget; it outputs a discrete action over
\{call agent $i$, HALT\}. Aggregation is equal-weight over the called
agents at termination. The policy is trained with PPO on a mixed
validation pool across datasets.

% [Result preview]
Our empirical analysis quantifies three facts:
(i) per-window oracle CRPS exceeds best-single by $+5.94\%$ on average
on ETT-small, providing a real upper bound for routing;
(ii) Sequential Refinement extracts a non-trivial fraction of this
headroom on a same-distribution diagnostic split, while a val/test
distribution shift in standard ETT splits caps the achievable gain;
(iii) the learned policy attains near-nominal coverage on 3/4 datasets
at $\approx 30\%$ fewer calls than the full ensemble, and beats a
split-conformal single-agent baseline on the coverage--cost frontier.

% [Limitations preview]
We are explicit about the empirical ceiling of routing on these
benchmarks at H=96; an extensive ablation on the per-window
oracle gap (\S\ref{sec:ceiling}) documents when routing can and cannot
help. We discuss extensions to longer horizons and broader dataset
suites in \S\ref{sec:discussion}.

\section{Related Work}
\label{sec:related}

% Foundation TSF models
% Conformal prediction for TS
% Multi-agent LLM debate (Puppeteer / Router-R1)
% Multi-model ensembling for TSF
% Cost-aware routing (FrugalGPT etc.)

\section{Method}
\label{sec:method}

\subsection{Panel of forecasters}
\label{sec:panel}

We use four agents covering distinct mechanisms:
\begin{itemize}
\item \textbf{ARIMA(2,1,1)}: statistical anchor.
\item \textbf{Chronos-Bolt-base} (T5 encoder-decoder, value tokenization).
\item \textbf{Chronos-Bolt-tiny}: smaller variant for compute--accuracy variation.
\item \textbf{Moirai-1.1-R-base} (encoder-decoder with masked attention, sample head).
\end{itemize}

Each agent emits a quantile forecast at levels $\{0.1, 0.25, 0.5, 0.75, 0.9\}$
over a horizon $H=96$.

\subsection{MDP formulation}
\label{sec:mdp}

\paragraph{State.} Per-window features (per-window normalized: mean, std,
min, max, range, $q_{25}$, $q_{75}$, skew, lag-1 autocorr, trend slope),
the current ensemble's flattened quantile array, the central interval
width, a binary mask of already-called agents, and the per-episode call
count. We optionally augment with a one-hot dataset identifier (Appendix~A
analyzes the impact).

\paragraph{Action.} Discrete in $\{0, 1, \ldots, N_{\text{agents}}\}$: actions
$1$--$N$ invoke the corresponding agent, action $0$ halts. Already-called
agents are masked at the policy logits.

\paragraph{Reward.}
\begin{align}
r_t = \begin{cases}
-\lambda_{\text{cost}} \cdot c(\text{agent}_i) & \text{if } a_t = \text{call agent } i, \\
-\text{CRPS}(F_{\text{ensemble}}, y) & \text{if } a_t = \text{HALT or budget exhausted}, \\
-0.1 & \text{if } a_t \text{ is invalid (re-call)}.
\end{cases}
\end{align}

\paragraph{Policy.} Two-layer MLP with $128$ hidden units, softmax over
$N_{\text{agents}}+1$ actions and a scalar value head.

\subsection{Training}
PPO with clipped surrogate objective, GAE, action masking on logits.
Hyperparameters listed in Appendix~B.

\section{Experiments}
\label{sec:experiments}

\subsection{Setup}

We use the ETT-small suite at lookback $L=96$ and horizon $H=96$. Splits
follow the Informer convention. Metrics: MSE/MAE on the $z$-scored scale,
quantile-CRPS, empirical coverage of the central 80\% interval, and mean
agent-calls per forecast.

\subsection{Per-agent baselines and oracle headroom}
\label{sec:headroom}

Table~\ref{tab:headroom} reports per-agent CRPS and coverage, the
oracle per-window CRPS, and the headroom of the latter over the
best single agent.

\begin{table}[h]
  \centering
  \caption{Per-agent and oracle CRPS on ETT-small ($H=96$, 256 test windows).}
  \label{tab:headroom}
  \begin{tabular}{lrrrr}
  \toprule
    & ETTh1 & ETTh2 & ETTm1 & ETTm2 \\
  \midrule
  ARIMA          & 0.204 & 0.382 & 0.123 & 0.588 \\
  Chronos-base   & 0.136 & 0.179 & 0.113 & 0.270 \\
  Chronos-tiny   & 0.134 & 0.179 & 0.103 & 0.305 \\
  Moirai         & 0.148 & 0.172 & 0.142 & 0.387 \\
  \midrule
  Best single    & 0.134 & 0.172 & 0.103 & 0.270 \\
  Oracle (per-w) & 0.123 & 0.156 & 0.097 & 0.268 \\
  Headroom (\%)  & +7.7  & +9.3  & +5.9  & +0.9  \\
  \bottomrule
  \end{tabular}
\end{table}

\subsection{Comparison: RL routing vs.\ conformal calibration vs.\ ensembles}
\label{sec:comparison}

Table~\ref{tab:headline} reports the headline comparison. Two findings are
robust across datasets:

\begin{table}[h]
  \centering
  \caption{Headline comparison on ETT-small ($H=96$, 256 test windows).
    Conformal calibration uses 256 val windows (4 stratified 64-window blocks).
    RL is the trained orchestrator from Section~\ref{sec:method}.
    $\Delta_{cov}$ is the absolute deviation of $cov_{80}$ from nominal $0.80$.}
  \label{tab:headline}
  \small
  \begin{tabular}{lrrrrrr}
  \toprule
    & \multicolumn{2}{c}{ETTh1} & \multicolumn{2}{c}{ETTh2} & \multicolumn{2}{c}{ETTm1, ETTm2 (avg)} \\
    Method & CRPS & $\Delta_{cov}$ & CRPS & $\Delta_{cov}$ & CRPS & $\Delta_{cov}$ \\
  \midrule
  Best naive single             & 0.134 & 0.081 & 0.172 & 0.012 & 0.187 & 0.067 \\
  Best conformal-calibrated     & 0.146 & 0.164 & 0.190 & 0.154 & 0.188 & 0.107 \\
  Uniform (4 agents)            & 0.143 & 0.088 & 0.205 & 0.012 & 0.242 & 0.012 \\
  Conformal-calibrated uniform  & 0.142 & 0.066 & 0.214 & 0.114 & 0.243 & 0.066 \\
  \textbf{RL orchestrator}      & 0.148 & 0.080 & 0.198 & 0.003 & 0.234 & 0.005 \\
  \bottomrule
  \end{tabular}
\end{table}

\paragraph{Finding 1: Split conformal overcorrects.} On three of four ETT
datasets, conformal-calibrated single agents push $cov_{80}$ above $0.95$
(see Appendix~A for per-dataset numbers). This is consistent with the val/
test distribution shift documented in Section~\ref{sec:ceiling}: the
conformity scores estimated on val are systematically pessimistic for test.
The resulting interval inflation costs CRPS even though average coverage
appears ``fixed''.

\paragraph{Finding 2: Multi-agent ensembling provides robust calibration
without conformal's overcorrection.} The uniform ensemble of four
heterogeneous agents lands cov$_{80}$ within $0.03$ of nominal on three
datasets without any val-based correction. The RL orchestrator does the
same with $\approx 30\%$ fewer agent calls (mean $2.7$ vs.\ $4$) and a
better calibration profile (mean $\Delta_{cov}=0.022$ vs.\ $0.030$
for uniform).

\subsection{Same-distribution diagnostic}
\label{sec:ceiling}

To distinguish method-ceiling from val/test distribution-shift, we split
each dataset's test set 80/20 randomly, train one policy on the 80\%
mixture across datasets, and evaluate on the 20\% holdouts. Mean RL CRPS
delta vs.\ best-single is $-1.36\%$, indicating that the method has a
ceiling at $H=96$ on ETT even under distribution-matched training.

\section{Discussion and Limitations}
\label{sec:discussion}

\paragraph{Empirical ceiling.} Per-window routing extracts only a small
fraction of the oracle headroom in our experiments. The bottleneck is
\emph{conditional} information: features that distinguish which agent
will win on a given window. Replacing hand-crafted statistics with a
learned encoder (Conv1D on raw lookback) is a natural extension.

\paragraph{Single benchmark.} Results are restricted to ETT-small at $H=96$.
Behavior at longer horizons (Chronos warns at $H>64$) and across more
heterogeneous datasets (Weather, Electricity, Traffic, M-competition,
GIFT-Eval) is an open question we plan to address.

\paragraph{Differentiating from concurrent work.} Puppeteer~\cite{wei2025puppeteer}
and Router-R1~\cite{tang2025routerr1} train RL orchestrators for text-domain
MDPs (reasoning, QA). Their predictions are tokens; ours are predictive
distributions over a horizon. Their state has no calibration features
or interval widths; ours does. Their halt rule is a learned token /
terminator agent; ours is a learned discrete HALT conditioned on the
ensemble's running quantile state.

\bibliographystyle{plain}
\bibliography{references}
\end{document}
